Visual Analysis of Ligand Trajectories in Molecular Dynamics

### 2018-09-28 paper.tex
Updates abstract base ond the low level comments from Helwig (TODO: implement the high level comments as well).

%% Abstract section.
\abstract{
Exploration of protein reactivity with other molecules plays a crucial role in many biochemical research disciplines, including protein engineering and drug design.
In many cases, a small molecule (ligand) reacts with the protein in its deeply buried active site. 
As the ligand has to be transported to this site from the protein outer environment, it is crucial to study trajectories, which the ligand can follow.
There are several in-silico methods, enabling to calculate potential trajectories in a protein model.
When these methods are applied in connection with molecular dynamics simulation, usually large ensembles of trajectories are produced.
These trajectories have to be further explored by the biochemists who are trying to assess their relevance and gain insight into the protein function.
During this process, many parameters of the trajectories have to be taken into account.
This makes the exploration very demanding as the biochemists have to analyze the multidimensional data about trajectories.
Without proper visual support, this process is very complicated and time demanding.
In this paper, we introduce a novel tool for visual analysis of ligand trajectories and their various properties.
The tool has been designed in tight collaboration with the domain experts in order to address their needs as much as possible.
It is composed of several 2D and 3D linked views, enabling interactive exploration and filtering of trajectories.
%The tool consists of a set of 2D views and the 3D view.
It utilizes the concept of small multiples for communicating the properties of the trajectories.
Through a variety of interaction techniques, the system enables users to create and store selections that can be compared and evaluated later.
In both 2D and 3D views, users can either observer the individual trajectories or choose to aggregate the information into a functional boxplot or a density visualization.
%It utilizes the concept of a scatterplot matrix for communicating different properties of the trajectories.
The usefulness of the tool is demonstrated in two case studies, conducted by the collaborating protein engineers.
} % end of abstract

>>>>

%% Abstract section.
\abstract{
Exploration of protein reactions with other molecules plays a crucial role in biochemical research, including protein engineering and drug design.
In many cases, a small molecule (ligand) reacts with the protein at a deeply buried active site. 
As the ligand has to get to this site from the outer environment, it is crucial to study trajectories of ligand motion.
There are several in-silico methods to calculate potential trajectories.
When these methods are applied in connection with molecular dynamics simulation, usually large ensembles of trajectories are produced.
These trajectories have to be explored by biochemists who are trying to assess their relevance and gain insight into the protein function.
During this process, many parameters of the trajectories have to be taken into account.
This makes the exploration very demanding as the biochemists have to analyze the multidimensional data about trajectories.
Without proper visual support, this process is very complicated and time demanding.
In this paper, we introduce a novel tool for the interactive visual exploration and analysis of ligand trajectories and their various properties.
The tool has been designed in tight collaboration with domain experts in order to address their needs as much as possible.
It is composed of several linked 2D and 3D views, enabling the interactive exploration and filtering of trajectories.
%The tool consists of a set of 2D views and the 3D view.
Small multiples are used for communicating the properties of the trajectories.
Through a variety of interaction techniques, the system enables users to create and store selections that can be compared and evaluated later.
In both 2D and 3D views, users can either observer the individual trajectories or choose to aggregate the information into a functional boxplot or a density visualization.
%It utilizes the concept of a scatterplot matrix for communicating different properties of the trajectories.
The usefulness of the tool is demonstrated in two case studies, conducted by the collaborating protein engineers.
} % end of abstract
